---
title: "GIS workshop"
author: "Darya Vanichkina"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rgdal)
library(raster)
library(tidyverse)
GDALinfo("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
DSM_HARV <- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
DSM_HARV
summary(DSM_HARV)
summary(DSM_HARV, maxsamp = ncell(DSM_HARV))
DSM_HARV_df <- as.data.frame(DSM_HARV, xy = TRUE)
str(DSM_HARV_df)
```

Plot data using ggplot

```{r}
ggplot() + geom_raster(data = DSM_HARV_df, aes(x = x, y = y, fill = HARV_dsmCrop)) + scale_fill_viridis_c() + coord_quickmap() + theme_bw()
```
CRS?

```{r}
crs(DSM_HARV)
```


Min and max of this raster?

```{r}
DSM_HARV
minValue(DSM_HARV)
maxValue(DSM_HARV)
# if it wasn't known
DSM_HARV <- setMinMax(DSM_HARV)
```

```{r}
GDALinfo("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")
nlayers(DSM_HARV)
```
Missing data?

Bad data values
```{r}
ggplot() + geom_histogram(data = DSM_HARV_df, aes(x = HARV_dsmCrop), bins = 50) + theme_bw()
```

Challenge:
Explore HARV_DSMhill.tif

1. Does this file have the same CRS? Yes
2. What is the NoDataValue? -9999
3. Resolution?  Same as DSM_HARV? 1m x 1 m ; yes
4. How large would a 5x5 pixel be? 5m x 5m 
5. Single or multiband raster? single
6. Bad values?

```{r}
GDALinfo("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_DSMhill.tif")
DSM_hill_HARV <- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_DSMhill.tif")
as.character(crs(DSM_hill_HARV)) == as.character(crs(DSM_HARV))

DSM_hill_HARV %>% as.data.frame(xy=TRUE) %>% ggplot() + geom_histogram( aes(x = HARV_DSMhill), bins = 50) + theme_bw()
```

```{r}
DSM_HARV_df<-DSM_HARV_df %>% 
  mutate(fct_elevation=cut(HARV_dsmCrop, breaks=3))
ggplot()+
  geom_bar(data=DSM_HARV_df,aes(fct_elevation))
```

```{r}
unique(DSM_HARV_df$fct_elevation)
```

```{r}
DSM_HARV_df %>% group_by(fct_elevation) %>%
  summarize(counts=n())
```

```{r}
custom_bins<- c(300,350,400, 450)
DSM_HARV_df<-DSM_HARV_df %>%
  mutate(fct_elevation_2=cut(HARV_dsmCrop,breaks=custom_bins))
unique(DSM_HARV_df$fct_elevation_2)
ggplot()+
  geom_bar(data=DSM_HARV_df,aes(fct_elevation_2))

```

```{r}
DSM_HARV_df %>%
  group_by(fct_elevation_2) %>%
  summarize(counts=n())
```
```{r}
ggplot()+
  geom_raster(data = DSM_HARV_df,aes(x=x,y=y, fill= fct_elevation_2))+
  coord_quickmap()
```

```{r}
terrain.colors(3)
```

```{r}
ggplot()+
  geom_raster(data = DSM_HARV_df,aes(x=x,y=y, fill= fct_elevation_2))+
  scale_fill_manual(values=terrain.colors(3))
  coord_quickmap()
```

```{r}
my_col<-terrain.colors(3)
ggplot()+
  geom_raster(data = DSM_HARV_df,aes(x=x,y=y, fill= fct_elevation_2))+
  scale_fill_manual(values=my_col, name="Elevation")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+

  coord_quickmap()

```
challenge
Create a plot of the Harvard Forest Digital Surface Model (DSM) that has:

Six classified ranges of values (break points) that are evenly divided among the range of pixel values.
Axis labels.
A plot title.
```{r}
DSM_HARV_df<- DSM_HARV_df %>% 
  mutate(fct_elevation_6=cut(HARV_dsmCrop, breaks=6))

my_col<- terrain.colors(6)
ggplot()+
  geom_raster(data = DSM_HARV_df,aes(x=x,y=y, fill= fct_elevation_6))+
  scale_fill_manual(values=my_col, name="Elevation")+
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())+

  coord_quickmap()

```
```{r}
DSM_hill_HARV <-
  raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_DSMhill.tif")

DSM_hill_HARV
```
```{r}
DSM_hill_HARV_df<- as.data.frame(DSM_hill_HARV,xy=TRUE)
```
```{r}
str(DSM_hill_HARV_df)
```
```{r}
ggplot()+ geom_raster(data=DSM_hill_HARV_df, 
                      aes(x=x, y=y, alpha=HARV_DSMhill))+
                            scale_alpha(range=c(0.15,0.65),guide="none")+
                            coord_quickmap()
                            
```

```{r}
ggplot()+
  geom_raster(data = DSM_HARV_df,aes(x=x,y=y, fill=HARV_dsmCrop))+

geom_raster(data=DSM_hill_HARV_df, 
                      aes(x=x, y=y, alpha=HARV_DSMhill))+
  scale_fill_viridis_c()+
  scale_alpha(range = c(0.15,0.65),guide="none")+
  ggtitle("Elevation with hillshade")+
  coord_quickmap()
  
```
```{r}
# import DSM data
DSM_SJER <- raster("data/NEON-DS-Airborne-Remote-Sensing/SJER/DSM/SJER_dsmCrop.tif")
# convert to df for plotting
DSM_SJER_df<-as.data.frame(DSM_SJER, xy=TRUE)
#Import DMS hillShade
DSM_hill_SJER<- raster("data/NEON-DS-Airborne-Remote-Sensing/SJER/DSM/SJER_dsmHill.tif")
# conver to DF for plotting

DSM_hill_SJER_df<-as.data.frame(DSM_hill_SJER, xy=TRUE)
# build plot
ggplot()+
  geom_raster(data = DSM_SJER_df, 
              aes(x=x, y=y, 
                  fill=SJER_dsmCrop,
                  alpha=0.8)
              )+
  geom_raster(data = DSM_hill_SJER_df,
              aes(x=x,y=y, 
                  alpha= SJER_dsmHill)
              )+
  scale_fill_viridis_c()+
  guides(fill=guide_colorbar())+scale_alpha(range=c(0.4, 0.7), guide="none")
  
  # remove gray bachgropund nad grid lines
  # theme_bw()+
  # theme(panel.grid.major = element_blank(),
  # panel.grid.m
```

```{r}

DTM_HARV <- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/DTM/HARV_dtmCrop.tif")

DTM_hill_HARV<-  raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/DTM/HARV_DTMhill_WGS84.tif")

# df
DTM_HARV_df<- as.data.frame(DTM_HARV, xy=TRUE)
DTM_hill_HARV_df<-as.data.frame(DTM_hill_HARV, xy=TRUE)


```


```{r}
crs(DTM_HARV )
```
```{r}
crs(DTM_hill_HARV)
```
```{r}
# projectRaster
DTM_hill_UTMZ18n_HAVR<- projectRaster(DTM_hill_HARV, crs = crs(DTM_HARV))
```

```{r}
crs(DTM_hill_UTMZ18n_HAVR)
```
```{r}
extent(DTM_hill_UTMZ18n_HAVR)
```

```{r}
extent(DTM_hill_HARV)
```

```{r}
res(DTM_hill_UTMZ18n_HAVR)
```

```{r}
res(DTM_hill_HARV)
```
```{r}
DTM_hill_UTMZ18N_HAVR <- projectRaster(DTM_hill_HARV, crs = crs(DTM_HARV), res=c(1,1))
```
```{r}
res(DTM_hill_UTMZ18N_HAVR )
```

```{r}
crs(DTM_hill_UTMZ18N_HAVR )
```
```{r}
DTM_hill_HARV_2_df<- as.data.frame(DTM_hill_UTMZ18N_HAVR, xy=TRUE)
```

```{r}
ggplot()+
  geom_raster(data = DTM_HARV_df,
              aes(x=x, y=y, 
                  fill = HARV_d
                  tmCrop))+
geom_raster(data = DTM_hill_HARV_2_df,
            aes(x=x,y=y,
                alpha=HARV_DTMhill_WGS84))+
  scale_fill_gradientn(name="Elevation", colors=terrain.colors(10))+
  coord_quickmap()
```
calculate the difference between two raster
1- subtracting
2-overlay()

```{r}
CHM_HARV <- DSM_HARV -DTM_HARV
#df
CHM_HARV_df<- as.data.frame(CHM_HARV ,xy=TRUE)

```

```{r}

ggplot()+
  geom_raster(data= CHM_HARV_df, 
              aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(name="Canpoy Height", colors = terrain.colors(10))+
  coord_quickmap()
```
```{r}
ggplot(CHM_HARV_df)+geom_histogram(aes(layer))
```

```{r}
CHM_ov_HARV <- overlay(DSM_HARV,DTM_HARV,
                       fun= function(r1,r2) {return(r1-r2)})
```
```{r}
# dataframe
CHM_ov_HARV_df<- as.data.frame(CHM_ov_HARV, xy=TRUE)
```

```{r}
ggplot()+
  geom_raster(data= CHM_ov_HARV_df, 
              aes(x=x, y=y, fill=layer))+
  scale_fill_gradientn(name="Canpoy Height", colors = terrain.colors(10))+
  coord_quickmap()
```
```{r}
# export a geoTiff
writeRaster(CHM_ov_HARV,"CHM_HARV.tiff", format="GTiff", overwrite=TRUE, NAflag=-9999)
```
 ## git
```{r}
RGB_band1_HARV <- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif")
```
```{r}
RGB_band1_HARV_df  <- as.data.frame(RGB_band1_HARV, xy = TRUE)

```

```{r}
ggplot() +
  geom_raster(data = RGB_band1_HARV_df,
              aes(x = x, y = y, alpha = HARV_RGB_Ortho)) + 
  coord_quickmap()
```

```{r}
RGB_band1_HARV
```

```{r}
nbands(RGB_band1_HARV)
nlayers(RGB_band1_HARV)
```
```{r}
RGB_band2_HARV <-  raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif", band = 2)

```

```{r}
RGB_band2_HARV_df <- as.data.frame(RGB_band2_HARV, xy = TRUE)

```
```{r}
RGB_stack_HARV <- stack("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif")
```

```{r}
RGB_stack_HARV

```

```{r}
RGB_stack_HARV@layers
```

```{r}
RGB_stack_HARV[[2]]

```
```{r}

RGB_band1_HARV<- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif")
RGB_band1_HARV_df<- as.data.frame(RGB_band1_HARV,xy=TRUE)
ggplot()+
  geom_raster(data = RGB_band1_HARV_df, aes(x=x, y=y, alpha=HARV_RGB_Ortho))+
  coord_quickmap()
```

```{r}
RGB_band1_HARV
```
```{r}

RGB_band2_HARV<- raster("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif", band=2)
RGB_band2_HARV_df<- as.data.frame(RGB_band2_HARV,xy=TRUE)
ggplot()+
  geom_raster(data = RGB_band2_HARV_df, aes(x=x, y=y, alpha=HARV_RGB_Ortho))+
  coord_quickmap()
```
```{r}
  RGB_stack_HARV<-stack("data/NEON-DS-Airborne-Remote-Sensing/HARV/RGB_Imagery/HARV_RGB_Ortho.tif")

 RGB_stack_HARV

 RGB_stack_HARV@layers

RGB_stack_HARV[[2]]
```
```{r}
# plotting
RGB_stack_HARV_df<-as.data.frame(RGB_stack_HARV,xy=TRUE)
```

```{r}
str(RGB_band1_HARV_df)
```

```{r}
str(RGB_stack_HARV_df)
```
```{r}
ggplot()+
  geom_histogram(data=RGB_stack_HARV_df, aes(HARV_RGB_Ortho.1))
```
```{r}
ggplot()+
  geom_raster(data=RGB_stack_HARV_df, aes(x=x, y=y, alpha=HARV_RGB_Ortho.2))+
  coord_quickmap()
```

```{r}
 plotRGB(RGB_stack_HARV, r=1, g=2, b=3)
```
```{r}
plotRGB(RGB_stack_HARV, r=1, g=2, b=3, scale=800,  stretch="hist")
```

```{r}
plotRGB(RGB_stack_HARV, r=1, g=2, b=3, scale=800,  stretch="lin")
```
```{r}
RGB_brick_HARV<- brick(RGB_stack_HARV)
```
```{r}
object.size(RGB_stack_HARV)
```
```{r}
object.size(RGB_brick_HARV)
```
```{r}
methods(class= class(RGB_stack_HARV))
methods(class=class(RGB_stack_HARV[1]))

```

```{r}
library(sf)
aoi_boundary_HARV <- st_read("data/NEON-DS-Site-Layout-Files/HARV/HarClip_UTMZ18.shp")
```

```{r}
st_geometry_type(aoi_boundary_HARV)
levels(st_geometry_type(aoi_boundary_HARV))
st_crs(aoi_boundary_HARV)
st_bbox(aoi_boundary_HARV)
ncol(aoi_boundary_HARV)
ggplot() + geom_sf(data = aoi_boundary_HARV, size = 3, colour = "black", fill = "cyan1") +
  ggtitle("AOI boundary plot") +
  coord_sf()


```

```{r}
# import HARV roads and HARV tower UTM18N layers as two shapefiles
# lines_HARV
# HARVtower_UTM18N

lines_HARV <- st_read("data/NEON-DS-Site-Layout-Files/HARV/HARV_roads.shp")
point_HARV <- st_read("data/NEON-DS-Site-Layout-Files/HARV/HARVtower_UTM18N.shp")
# Q1: what type object?
# Q2: what is the CRS and extent?
# Q3: Points/lines/polygons?
# Q4: How many spatial objects in each file?
```



```{r}
names(lines_HARV)
class(lines_HARV)
head(lines_HARV)
summary(lines_HARV)
```


```{r}
levels(lines_HARV$TYPE)
table(lines_HARV$TYPE)
head(lines_HARV)

lines_HARV %>% filter(TYPE == "footpath") %>% ggplot() + geom_sf(aes(color = as.factor(OBJECTID_1) ), size = 1.5) +ggtitle("NEON HARV field site", subtitle = "Footpath") + coord_sf() + theme_bw() + labs(color = "FootpathID")
#  Plot only the boardwalk (no new object!)
lines_HARV %>% filter(TYPE == "boardwalk") %>% ggplot() + geom_sf() + ggtitle("HARV", subtitle = "Boardwalk") + coord_sf() + theme_bw()
# Plot only the stone walls (no new object)

lines_HARV %>% filter(TYPE == "stone wall") %>% ggplot() + geom_sf(aes(col = as.factor(OBJECTID_1))) +coord_sf() + theme_bw()
```


```{r}
lines_HARV %>% ggplot() + geom_sf(aes(col = as.factor(OBJECTID_1), size = TYPE)) + labs(col= "ObjectID") + scale_size_manual(values = c(1,2,3,4)) + coord_sf() + theme_bw() 
```

```{r}
# install.packages("tmap")
library(tmap)
tm_shape(lines_HARV) + tm_lines("TYPE", palette = "Set2")
```

Challenge: plot lines by attribute (using ggplot)
Create a plot which emphasises only roads where bicycles and horses are allowed. Make lines where they ARE allowed thinner and in a bright color. Add a title and legend to the plot. Make roads where they're not allowed black. 

```{r}
BikeHorseAllowed <- lines_HARV %>% filter(BicyclesHo == "Bicycles and Horses Allowed")
ggplot() + 
  geom_sf(data = lines_HARV) +
  geom_sf(data = BikeHorseAllowed, aes(col = BicyclesHo),  size = 2, col = "magenta")

lines_HARV %>% filter(BicyclesHo == "Bicycles and Horses Allowed") %>% ggplot() + geom_sf(aes(col = BicyclesHo),  size = 2, col = "magenta") + geom_sf(data = lines_HARV)

```



```{r}
# state boundaries US
state_boundary_US <- st_read("data/NEON-DS-Site-Layout-Files/US-Boundary-Layers/US-State-Boundaries-Census-2014.shp")
# Create a map of US State Boundaries coloured by region. 

state_boundary_US %>% ggplot() + geom_sf(aes(col= region)) + theme_minimal() + coord_sf()

tm_shape(state_boundary_US) + tm_polygons("region", palette = "Set3")
tm_shape(state_boundary_US) + tm_borders()
tm_shape(state_boundary_US) + tm_fill("region", palette = "Set3")

plot(state_boundary_US["region"])
```

```{r}
#point_HARV <- whatvert
ggplot() + geom_sf(data = aoi_boundary_HARV, fill = "pink", col = "pink") +
  geom_sf(data = point_HARV, aes(fill = Sub_Type), shape = 15) +
  geom_sf(data = lines_HARV, aes(col = TYPE), show.legend = "line", size = 1)
```
Plot raster AND vector together:

- CHM raster from yesterday
- lines_HARV
- aoi_boundary_HARV
- point_HARV

```{r}
ggplot() + geom_raster(data = CHM_HARV_df, aes(x = x, y = y, fill = layer))  + 
  geom_sf(data = lines_HARV, color = "black") +
  geom_sf(data = aoi_boundary_HARV, colour = "grey20", size = 1) + 
  geom_sf(data = point_HARV, pch = 8)

tm_shape(CHM_HARV) + tm_raster(alpha = 0.8) + tm_shape(lines_HARV) + tm_lines() + tm_shape(aoi_boundary_HARV) + tm_borders("blue") + tm_shape(point_HARV) + tm_symbols(size = 0.25, col = "red")
```

